---
title: "Brain DIA Proteomics Analysis Pipeline"
author: "Ravi S Pandey"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, error = FALSE, echo = FALSE)
```

---

## ⚙️ USER CONFIGURATION — Edit This Section Before Running

```{r user_config, echo=TRUE}
# ── PATHS ─────────────────────────────────────────────────────────────────────
data_dir       <- "../data/example"           # Directory containing input data files
result_dir     <- "../results/"        # Directory to save output results

# Input file names (place these in data_dir)
abundance_file  <- "example_abundance.csv"   # Normalized protein abundance matrix
traits_file     <- "example_traits.csv"     # Sample metadata / traits file

# ── STUDY DESIGN ──────────────────────────────────────────────────────────────
# Name of the control genotype (must match values in traits file exactly)
control_genotype <- "Control_Genotype_Name"

# Names of case/treatment genotypes (must match values in traits file exactly)
case_genotypes <- c(
  "Case_Genotype_1",
  "Case_Genotype_2",
  "Case_Genotype_3"
)

# Age groups in your study (numeric values as they appear in traits file)
age_groups <- c(4, 12)   # e.g., c(4, 12) for 4-month and 12-month timepoints

# Sex column values in traits file
sex_levels <- c("Male", "Female")

# ── SPECIES ───────────────────────────────────────────────────────────────────
# Organism database for annotation (installed Bioconductor package)
# Common options:
#   Mouse: "org.Mm.eg.db"
#   Human: "org.Hs.eg.db"
#   Rat:   "org.Rn.eg.db"
organism_db  <- "org.Mm.eg.db"
kegg_organism <- "mmu"    # KEGG organism code: "mmu"=mouse, "hsa"=human, "rno"=rat

# ── STATISTICAL THRESHOLDS ────────────────────────────────────────────────────
pval_cutoff   <- 0.05     # Adjusted p-value cutoff for significance
lfc_cutoff    <- 0        # Log2 fold-change cutoff (0 = any direction)
top_pathways  <- 10       # Number of top pathways to display in enrichment plots

# ── COLUMN NAME MAPPING ───────────────────────────────────────────────────────
# Column names in your traits/metadata file
col_sample_id  <- "X"          # Sample ID column
col_sex        <- "Sex"        # Sex column
col_genotype   <- "Genotype"   # Genotype column
col_age        <- "Age"        # Age column

# ── PROTEIN ID FORMAT ─────────────────────────────────────────────────────────
# Does the ProteinID column contain pipe-delimited entries (Protein|Peptide)?
# Set TRUE if raw DIA data has multiple peptide entries per protein
has_peptide_suffix <- TRUE

# ── OPTIONAL: CO-EXPRESSION MODULE CORRELATION ────────────────────────────────
# Set to TRUE only if you have external reference module data
run_module_correlation <- FALSE
module_data_file <- ""    # Path to module reference data (.RData or .csv)

# ── OPTIONAL: BIOLOGICAL DOMAIN ANALYSIS ─────────────────────────────────────
# Set to TRUE only if you have biodomain annotation data
run_biodomain_analysis <- FALSE
biodomain_data_file <- ""  # Path to biodomain annotation data
```

---

# {.tabset .tabset-fade}

## Data Overview

### 1. Experimental Design

This document reports the processing of Brain DIA Proteomics data.

**Study Groups:**

- **Control:** `r control_genotype`
- **Cases:** `r paste(case_genotypes, collapse = ", ")`
- **Ages:** `r paste(age_groups, collapse = " and ")` months
- **Sex:** `r paste(sex_levels, collapse = " and ")`

**Analysis includes:**

- Protein-level QC and normalization assessment
- PCA for sample-level quality control
- Differential abundance analysis (ANOVA + Tukey HSD)
- KEGG pathway enrichment (ORA and GSEA)
- Visualization: volcano plots, upset plots, barplots

---

### 2. Data Loading & Initial QC

#### 2.1 Load Required Libraries

```{r library, echo=TRUE}
source("../R/Helper_Functions_Prot.R")
```

#### 2.2 Load & Format Metadata

```{r metadata}
# Load sample metadata
trait_data <- read.csv(file.path(data_dir, traits_file)) %>%
  dplyr::select(
    ID         = !!col_sample_id,
    Sex        = !!col_sex,
    Genotype   = !!col_genotype,
    Age        = !!col_age
  ) %>%
  # Keep only genotypes defined in config
  filter(Genotype %in% c(control_genotype, case_genotypes))

# Display sample summary table
study_summary_table <- trait_data %>%
  dplyr::count(Genotype, Age, Sex) %>%
  pivot_wider(names_from = Sex, values_from = n, values_fill = 0) %>%
  mutate(Age = factor(Age, levels = sort(age_groups))) %>%
  arrange(Genotype, Age) %>%
  gt() %>%
  tab_header(
    title    = md("**Study Population Summary**"),
    subtitle = "Brain Proteomics (DIA)"
  ) %>%
  cols_label(Age = "Age Group") %>%
  tab_style(
    style     = cell_fill(color = "gray96"),
    locations = cells_body(rows = Genotype == control_genotype)
  ) %>%
  grand_summary_rows(
    columns = intersect(sex_levels, names(.)),
    fns     = list(Total = ~sum(.)),
    fmt     = list(~fmt_number(., decimals = 0))
  ) %>%
  tab_options(table.width = pct(80), column_labels.font.weight = "bold")

study_summary_table
```

#### 2.3 Protein Abundance Matrix

##### 2.3.1 Load Normalized Abundance Data

> **Note on normalization:** The protein abundances should be pre-normalized before input. This pipeline assumes batch effects have been removed (e.g., via TAMPOR or similar). If raw data is provided, add a normalization step before proceeding. Proteins with >50% missing values per group are recommended to be excluded upstream.

```{r QC1}
# Load abundance data
abundance_raw <- read.csv(
  file.path(data_dir, abundance_file),
  header     = TRUE,
  row.names  = 1,
  check.names = FALSE
) %>%
  rownames_to_column(var = "ProteinID")

# Display dimensions
cat("Dimensions:", dim(abundance_raw)[1], "proteins x", dim(abundance_raw)[2] - 1, "samples\n")
abundance_raw[1:5, 1:6]
```

Check sample-level distribution with boxplot:

```{r QC_boxplot, fig.width=14, fig.height=5}
df_check <- abundance_raw %>% column_to_rownames(var = "ProteinID")
boxplot(df_check, las = 2, main = "Sample Distributions (Post-Normalization)",
        ylab = "Abundance", cex.axis = 0.6)
```

##### 2.3.2 Summarize to Protein Level

```{r QC2}
if (has_peptide_suffix) {
  # Remove peptide suffixes (pipe-delimited format: Protein|Peptide)
  expr_mat <- abundance_raw %>%
    separate(col = ProteinID, into = c("Proteins", "Peptides"), sep = "[|]") %>%
    filter(!(Proteins %in% "")) %>%
    dplyr::select(-"Peptides")
} else {
  expr_mat <- abundance_raw %>% rename(Proteins = ProteinID)
}

# Collapse duplicates by averaging
clean_abundance <- expr_mat %>%
  group_by(Proteins) %>%
  summarise(across(everything(), mean)) %>%
  mutate(Proteins = str_to_title(Proteins)) %>%
  column_to_rownames(var = "Proteins")

cat("Final protein matrix:", nrow(clean_abundance), "proteins x", ncol(clean_abundance), "samples\n")
clean_abundance[1:5, 1:5]
```

##### 2.3.3 Merge Traits with Abundance Data

```{r combine}
# Join traits with expression (long format)
abundance_with_traits <- trait_data %>%
  dplyr::select(animalName = ID, Genotype, Sex, Age) %>%
  left_join(
    clean_abundance %>%
      t() %>%
      as.data.frame() %>%
      rownames_to_column(var = "animalName"),
    by = "animalName"
  )

abundance_with_traits[1:5, 1:9]

# Save processed data
save(trait_data, clean_abundance, abundance_with_traits,
     file = file.path(data_dir, "Processed_DIA_Proteomics.RData"))
```

---

## PCA

Principal component analysis (PCA) using probabilistic PCA to handle missing values.

```{r PCA, fig.width=12, fig.height=6}
load(file.path(data_dir, "Processed_DIA_Proteomics.RData"))
library(pcaMethods)

# Run probabilistic PCA (handles missing values)
pc_obj    <- pca(t(clean_abundance), method = "ppca", nPcs = 2)
pca_scores <- as.data.frame(scores(pc_obj)) %>%
  rownames_to_column(var = "SampleID")

# Join with metadata
plot_df <- left_join(pca_scores, trait_data, by = c(SampleID = "ID"))

# Plot
ggplot(plot_df, aes(x = PC1, y = PC2, color = Genotype, shape = Sex)) +
  geom_point(size = 3) +
  geom_text(aes(label = Age), hjust = 0.5, vjust = 2.5, size = 3.5, show.legend = FALSE) +
  theme_minimal() +
  labs(title = "PCA — DIA Proteomics (PPCA)")
```

---

## Differential Analysis Results {.tabset .tabset-fade}

One-way ANOVA followed by Tukey HSD post-hoc tests to identify differentially abundant proteins. Analyses are stratified by sex and age.

```{r load_data, echo=TRUE}
load(file.path(data_dir, "Processed_DIA_Proteomics.RData"))
```

### ANOVA

#### 4.1 Group Definition & Comparison Setup

```{r group_definition}
# Build all Genotype-Sex-Age combinations
all_groups <- expand.grid(
  Genotype = c(control_genotype, case_genotypes),
  Sex      = sex_levels,
  Age      = age_groups,
  stringsAsFactors = FALSE
) %>%
  mutate(Group = paste0(Genotype, "_", Sex, "_", Age, "M"))

# Add Group column to data
df <- abundance_with_traits %>%
  mutate(
    Group = paste0(Genotype, "_", Sex, "_", Age, "M"),
    .after = "Age"
  ) %>%
  mutate(Group = factor(Group, levels = all_groups$Group))

df[1:5, 1:9]
```

#### 4.2 Run DE
```{r}
comparisons  <- build_comparisons(df, control_genotype, case_genotypes, sex_levels, age_groups)
anova_output <- run_anova_all_comparisons(df, comparisons)
DE_results.list <- anova_output$results_list
DE_results.df   <- anova_output$results_df
save(DE_results.list, DE_results.df,
     file = file.path(result_dir, "ANOVA_Results_DIA_Proteomics.Rdata"))
```

#### 4.3 Results Summary & Visualization

```{r load_results}
load(file.path(result_dir, "ANOVA_Results_DIA_Proteomics.Rdata"))
head(DE_results.df)
```

```{r DEP_table, fig.width=10, fig.height=8}
degs.up   <- map(DE_results.list, ~length(which(.x$padj < pval_cutoff & .x$diff > lfc_cutoff)))
degs.down <- map(DE_results.list, ~length(which(.x$padj < pval_cutoff & .x$diff < -lfc_cutoff)))

deg_summary <- data.frame(
  comparison = names(degs.up),
  Up         = as.integer(unlist(degs.up)),
  Down       = as.integer(unlist(degs.down))
)

deg_summary %>%
  mutate(Sex = ifelse(grepl("Female", comparison), "Female", "Male")) %>%
  gt() %>%
  tab_header(title = md(paste0("**DEPs at adjusted p < ", pval_cutoff, "**")))
```

```{r DEP_barplot, fig.width=10, fig.height=6}
data_plot <- deg_summary %>%
  mutate(Sex = ifelse(grepl("Female", comparison), "Female", "Male")) %>%
  pivot_longer(cols = c(Up, Down), names_to = "Direction", values_to = "DEPs")

ggplot(data_plot, aes(x = comparison, y = DEPs, fill = Direction)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", linewidth = 0.3) +
  facet_wrap(~Sex, scales = "free_y") +
  scale_fill_manual(values = c("Up" = "#D55E00", "Down" = "#0072B2")) +
  labs(title = "Differentially Abundant Proteins (DAPs)",
       subtitle = paste0("adjP < ", pval_cutoff, " | |Log2FC| > ", lfc_cutoff),
       x = "", y = "Number of DAPs") +
  theme_bw() +
  theme(axis.text.x  = element_text(angle = 45, hjust = 1, size = 9),
        plot.title   = element_text(hjust = 0.5, face = "bold"),
        legend.position = "top")
```

#### 4.5 Expression of Selected Proteins

Modify the protein list below to plot proteins of interest:

```{r protein_plots, fig.width=12, fig.height=7}
# ── USER: Change these protein names to your proteins of interest ──────────
proteins_of_interest_female <- c("App", "Apoe")
proteins_of_interest_male   <- c("App", "Mapt")
# ──────────────────────────────────────────────────────────────────────────

female_degs <- DE_results.df %>% filter(grepl("Female", model))
male_degs   <- DE_results.df %>% filter(grepl("Male",   model))

plot_protein_expression(female_degs, proteins_of_interest_female, "Females")
plot_protein_expression(male_degs,   proteins_of_interest_male,   "Males")
```

### Overlap of DAPs

```{r upset_female, fig.width=9, fig.height=5,message=TRUE}
significant_f <- female_degs %>% filter(padj < pval_cutoff)
safe_upset(significant_f, direction = "up",   bar_color = "#164B6E", x_label = "Upregulated Proteins")
safe_upset(significant_f, direction = "down", bar_color = "#85070C", x_label = "Downregulated Proteins")
```

```{r upset_male, fig.width=9, fig.height=5,message=TRUE}
significant_m <- male_degs %>% filter(padj < pval_cutoff)
safe_upset(significant_m, direction = "up",   bar_color = "#164B6E", x_label = "Upregulated Proteins")
safe_upset(significant_m, direction = "down", bar_color = "#85070C", x_label = "Downregulated Proteins")
```

### Volcano Plots

```{r volcano, fig.height=8, fig.width=9}
for (i in names(DE_results.list)) {
  df_v <- DE_results.list[[i]] %>%
    as.data.frame() %>%
    mutate(log2FoldChange = as.numeric(diff), pvalue = as.numeric(padj))

  print(
    EnhancedVolcano(df_v,
      x = "log2FoldChange", y = "pvalue",
      lab = df_v$Symbol, title = i, subtitle = "",
      FCcutoff = lfc_cutoff, pCutoff = pval_cutoff,
      xlim = c(min(df_v$log2FoldChange, na.rm = TRUE) - 1,
               max(df_v$log2FoldChange, na.rm = TRUE) + 1),
      ylim = c(0, max(-log10(df_v$pvalue), na.rm = TRUE) + 2)
    ) +
      annotate("text",
        x = min(df_v$log2FoldChange) / 2,
        y = max(-log10(df_v$pvalue), na.rm = TRUE) + 1.5,
        size = 5,
        label = paste0("Down: ", nrow(filter(df_v, pvalue < pval_cutoff & log2FoldChange < 0)))
      ) +
      annotate("text",
        x = max(df_v$log2FoldChange) / 2,
        y = max(-log10(df_v$pvalue), na.rm = TRUE) + 1.5,
        size = 5,
        label = paste0("Up: ", nrow(filter(df_v, pvalue < pval_cutoff & log2FoldChange > 0)))
      )
  )
}
```

---

## Functional Analysis {.tabset .tabset-pills}

Over-representation analysis (ORA) and Gene Set Enrichment Analysis (GSEA) of differentially abundant proteins against KEGG pathways using `clusterProfiler`.

### KEGG Enrichment — Females

```{r kegg_female, fig.height=11, fig.width=11,message=TRUE}
dat_f <- female_degs %>%
  mutate(EntrezGene = AnnotationDbi::mapIds(
    get(organism_db),
    keys     = Symbol,
    column   = "ENTREZID",
    keytype  = "SYMBOL",
    multiVals = "first"
  ))

DEG_f <- dat_f %>%
  filter(padj < pval_cutoff) %>%
  arrange(padj) %>%
  dplyr::select(group = model, Symbol, EntrezGene, log2FoldChange = diff)

kegg.fn(unique(DEG_f$group), DEG_f, n = top_pathways)
```

### KEGG Enrichment — Males

```{r kegg_male, fig.height=11, fig.width=11,message=TRUE}
dat_m <- male_degs %>%
  mutate(EntrezGene = AnnotationDbi::mapIds(
    get(organism_db),
    keys     = Symbol,
    column   = "ENTREZID",
    keytype  = "SYMBOL",
    multiVals = "first"
  ))

DEG_m <- dat_m %>%
  filter(padj < pval_cutoff) %>%
  arrange(padj) %>%
  dplyr::select(group = model, Symbol, EntrezGene, log2FoldChange = diff)

kegg.fn(unique(DEG_m$group), DEG_m, n = top_pathways)
```

---

## Session Info

```{r session_info, echo=TRUE}
sessionInfo()
```
