---
title: "Human Translation Analysis — DIA Proteomics"
author: "Ravi S Pandey"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, error = FALSE, echo = FALSE)
```

> **Prerequisite:** Run `01_DIA_Proteomics_Pipeline.Rmd` first to generate
> `ANOVA_Results_DIA_Proteomics.Rdata`, which this notebook loads as input.

---

## ⚙️ USER CONFIGURATION — Edit This Section Before Running

```{r user_config, echo=TRUE}
# ── PATHS ─────────────────────────────────────────────────────────────────────
result_dir   <- "../results/"   # Must match result_dir in notebook 01
data_dir     <- "../data/"

# ── INPUT: ANOVA RESULTS FROM NOTEBOOK 01 ────────────────────────────────────
anova_results_file <- "ANOVA_Results_DIA_Proteomics.Rdata"

# ── HUMAN REFERENCE DATA ──────────────────────────────────────────────────────
# Co-expression module reference (e.g., Emory TMT proteomics modules)
# Expected columns: Symbol, Module, AD.Control
module_data_file <- ""          # Path to module reference .RData or .csv

# BioDomain reference files (from Agora / AMP-AD)
# biodom_path: .rds file with GO_ID → Biodomain → Subdomain mapping
# domain_colors_path: .csv with domain → color mapping
biodom_path         <- ""       # e.g., "data/annotated_biodomains.rds"
domain_colors_path  <- ""       # e.g., "data/domain_colors.csv"

# ── SPECIES / ORGANISM ────────────────────────────────────────────────────────
# Must match settings used in notebook 01
organism_db  <- "org.Mm.eg.db"   # "org.Mm.eg.db" = mouse, "org.Hs.eg.db" = human

# ── STATISTICAL THRESHOLDS ────────────────────────────────────────────────────
pval_cutoff   <- 0.05    # Significance cutoff used in notebook 01
lfc_cutoff    <- 0       # Log2FC cutoff used in notebook 01
gsea_seed     <- 1111    # Random seed for GSEA reproducibility

# ── ANALYSIS SWITCHES ─────────────────────────────────────────────────────────
# Set to TRUE only if you have the corresponding reference data files above
run_module_corr    <- FALSE   # Co-expression module correlation (ANOVA-based)
run_biodomain_ora  <- FALSE   # GO-ORA with BioDomain annotation
run_biodomain_gsea <- FALSE   # GO-GSEA with BioDomain annotation
run_lm_analysis    <- FALSE   # Linear modeling + human module/biodomain correlation

# ── PLOT DISPLAY OPTIONS ──────────────────────────────────────────────────────
top_biodomains  <- 5     # Number of top biodomains to highlight in sub-domain plots
nes_axis_limit  <- 3     # NES axis range for GSEA violin plots
top_modules     <- 10    # Number of top modules to display in correlation plots

# ── LINEAR MODEL CONFIGURATION ────────────────────────────────────────────────
# Only used when run_lm_analysis <- TRUE
#
# The LM tests the effect of each genotype while controlling for covariates.
# Formula: abundance ~ Genotype + Age + Sex  (or customize below)
lm_formula <- "~ Genotype + Age + Sex"

# Coefficient names to extract from the LM (must match factor levels in your data)
# These are the genotype contrasts you want to report — typically your case genotypes
# Example: c("(Intercept)", "GenotypeCaseA", "GenotypeCaseB", "Age", "SexMale")
# Leave NULL to auto-detect all coefficients
lm_coef_names <- NULL

# Factor reference level for Genotype (will be set as the baseline in the LM)
lm_reference_genotype <- control_genotype   # uses the value from above

# ── MODEL ORDERING (for plots) ────────────────────────────────────────────────
# Define the order in which models appear on plot axes.
# Leave NULL to use alphabetical order.
# Example: c("CaseA_Female_4M_vs_Control_Female_4M", ...)
order_model_female <- NULL
order_model_male   <- NULL

# Variant order for LM plots (coefficient names in desired display order)
# Leave NULL to use default order
order_variant_lm <- NULL

# ── LM VARIANT GROUPING ───────────────────────────────────────────────────────
# Maps each LM coefficient (Variant) to a display group for faceting in plots.
# Keys must exactly match coefficient names produced by your LM.
# Leave NULL to skip grouping (plots will not be faceted by group).
lm_variant_groups <- list(
  "SexMale"          = "Covariate",
  "Age"              = "Covariate",
  "GenotypeCaseA_4M" = "Genotype Effects",
  "GenotypeCaseB_4M" = "Genotype Effects"
)

# Factor order for groups in plots (controls facet order top to bottom)
lm_group_order <- c("Genotype Effects", "Covariate")
```

```{r load_organism_db, echo=TRUE}
# Load and resolve organism db — do this once here
library(organism_db, character.only = TRUE)
org_db_obj <- get(organism_db)
```

---

```{r source_helpers}
source("../R/Helper_Functions_Prot.R")
```

---

# {.tabset .tabset-fade}

## 1. Load Data

### 1.1 Load ANOVA Results from Notebook 01

```{r load_anova}
anova_file <- file.path(result_dir, anova_results_file)

if (!file.exists(anova_file))
  stop(
    "ANOVA results not found at: ", anova_file, "\n",
    "Please run 01_DIA_Proteomics_Pipeline.Rmd first."
  )

load(anova_file)

cat("Models loaded:", length(DE_results.list), "\n")
cat("Total DEG rows:", nrow(DE_results.df), "\n")
cat("Models:\n")
cat(paste0("  - ", names(DE_results.list), collapse = "\n"), "\n")
```

### 1.2 Stratify by Sex

```{r stratify}
female_degs <- DE_results.df %>% dplyr::filter(grepl("Female", model))
male_degs   <- DE_results.df %>% dplyr::filter(grepl("Male",   model))

cat("Female comparisons:", length(unique(female_degs$model)), "\n")
cat("Male comparisons:  ", length(unique(male_degs$model)),   "\n")
```

### 1.3 Add Entrez IDs

```{r add_entrez}
# Map gene symbols to Entrez IDs — done once outside mutate() to avoid scoping issues
entrez_map_f <- map_symbols_to_entrez(female_degs$Symbol, org_db_obj)
entrez_map_m <- map_symbols_to_entrez(male_degs$Symbol,   org_db_obj)

female_degs <- female_degs %>%
  dplyr::mutate(EntrezGene = entrez_map_f[Symbol])

male_degs <- male_degs %>%
  dplyr::mutate(EntrezGene = entrez_map_m[Symbol])

# Mapping summary
cat("Female — mapped:", sum(!is.na(female_degs$EntrezGene)),
    "| missing:", sum(is.na(female_degs$EntrezGene)), "\n")
cat("Male   — mapped:", sum(!is.na(male_degs$EntrezGene)),
    "| missing:", sum(is.na(male_degs$EntrezGene)), "\n")
```

### 1.4 Filter to Significant DEPs

```{r sig_deps}
sig_female <- female_degs %>%
  dplyr::filter(padj < pval_cutoff, abs(diff) > lfc_cutoff) %>%
  dplyr::arrange(padj)

sig_male <- male_degs %>%
  dplyr::filter(padj < pval_cutoff, abs(diff) > lfc_cutoff) %>%
  dplyr::arrange(padj)

cat("Significant DEPs — Female:", nrow(sig_female),
    "| Male:", nrow(sig_male), "\n")
```

---

## 2. Co-expression Module Correlation {.tabset .tabset-pills}

Pearson correlation between mouse model fold-changes and human AD brain
proteomics co-expression module signatures. Significant correlations (p < 0.05)
are highlighted with a black square overlay.

```{r module_check}
if (!run_module_corr) {
  cat("Module correlation skipped.\n",
      "Set run_module_corr <- TRUE and provide module_data_file to enable.\n")
  knitr::knit_exit()
}

if (!file.exists(module_data_file))
  stop("Module reference file not found at: ", module_data_file)

load_module_refs(module_data_file)
```

### 2.1 Female — Module Correlation

```{r module_corr_female, fig.width=12, fig.height=6, eval=run_module_corr}
cor_female <- run_module_correlation(female_degs, human_ref = module_ref)

if (!is.null(cor_female) && nrow(cor_female) > 0) {
  plot_module_corrplot(
    cor_data     = cor_female,
    order_model  = order_model_female,
    title        = "Human AD Module Correlation — Females"
  )
} else {
  message("No module correlations computed for females (insufficient overlapping genes).")
}
```

### 2.2 Male — Module Correlation

```{r module_corr_male, fig.width=12, fig.height=6, eval=run_module_corr}
cor_male <- run_module_correlation(male_degs, human_ref = module_ref)

if (!is.null(cor_male) && nrow(cor_male) > 0) {
  plot_module_corrplot(
    cor_data     = cor_male,
    order_model  = order_model_male,
    title        = "Human AD Module Correlation — Males"
  )
} else {
  message("No module correlations computed for males (insufficient overlapping genes).")
}
```

### 2.3 Scatter — Top Modules Female

```{r module_scatter_female, fig.width=12, fig.height=8, eval=run_module_corr}
# Identify top significant modules for female
if (!is.null(cor_female) && nrow(cor_female) > 0) {
  top_mods_f <- cor_female %>%
    dplyr::filter(significant) %>%
    dplyr::arrange(p_value) %>%
    dplyr::pull(Module) %>%
    unique() %>%
    head(top_modules)

  if (length(top_mods_f) > 0) {
    plot_module_scatter_correlations(
      deg_df       = female_degs,
      human_ref    = module_ref,
      mods_to_plot = top_mods_f,
      sex_label    = "Females"
    )
  } else {
    message("No significant modules to plot for females.")
  }
}
```

### 2.4 Scatter — Top Modules Male

```{r module_scatter_male, fig.width=12, fig.height=8, eval=run_module_corr}
if (!is.null(cor_male) && nrow(cor_male) > 0) {
  top_mods_m <- cor_male %>%
    dplyr::filter(significant) %>%
    dplyr::arrange(p_value) %>%
    dplyr::pull(Module) %>%
    unique() %>%
    head(top_modules)

  if (length(top_mods_m) > 0) {
    plot_module_scatter_correlations(
      deg_df       = male_degs,
      human_ref    = module_ref,
      mods_to_plot = top_mods_m,
      sex_label    = "Males"
    )
  } else {
    message("No significant modules to plot for males.")
  }
}
```

---

## 3. BioDomain — GO Over-Representation Analysis {.tabset .tabset-pills}

GO terms enriched in significantly differentially abundant proteins are mapped
to AD biological domains (Agora/AMP-AD BioDomain framework). This identifies
which disease-relevant biological processes are altered in each mouse model.

```{r biodomain_check}
if (!run_biodomain_ora) {
  cat("BioDomain ORA skipped.\n",
      "Set run_biodomain_ora <- TRUE and provide biodom_path and domain_colors_path to enable.\n")
}
```

```{r load_biodomain_refs, eval=run_biodomain_ora}
bd_refs <- load_biodomain_refs(biodom_path, domain_colors_path)
bio_ref <- bd_refs$biodom
lab_ref <- bd_refs$dom_lab
```

### 3.1 ORA — Females

```{r ora_female, fig.width=14, fig.height=10, eval=run_biodomain_ora}
if (nrow(sig_female) == 0) {
  message("ORA skipped for females: no significant DEPs at padj < ", pval_cutoff)
} else {
  enr_ora_f <- enrichORA_BD_Protein(
    dat         = sig_female,
    bio_ref     = bio_ref,
    lab_ref     = lab_ref,
    org_db      = org_db_obj,
    pval_cutoff = pval_cutoff
  )

  if (!is.null(enr_ora_f)) {
    enr.ORA.bd_plot(enr_ora_f)
  } else {
    message("No enriched GO terms found for females.")
  }
}
```

### 3.2 ORA — Males

```{r ora_male, fig.width=14, fig.height=10, eval=run_biodomain_ora}
if (nrow(sig_male) == 0) {
  message("ORA skipped for males: no significant DEPs at padj < ", pval_cutoff)
} else {
  enr_ora_m <- enrichORA_BD_Protein(
    dat         = sig_male,
    bio_ref     = bio_ref,
    lab_ref     = lab_ref,
    org_db      = org_db_obj,
    pval_cutoff = pval_cutoff
  )

  if (!is.null(enr_ora_m)) {
    enr.ORA.bd_plot(enr_ora_m)
  } else {
    message("No enriched GO terms found for males.")
  }
}
```

### 3.3 Sub-BioDomain ORA — Females

```{r ora_sub_female, fig.width=14, fig.height=16, eval=run_biodomain_ora}
if (!is.null(enr_ora_f) && nrow(enr_ora_f) > 0) {

  # Select top biodomains by number of enriched terms
  top_bd_f <- enr_ora_f %>%
    dplyr::filter(Biodomain != "none") %>%
    dplyr::count(Biodomain, sort = TRUE) %>%
    head(top_biodomains) %>%
    dplyr::pull(Biodomain)

  sub_ora_f <- enr_ora_f %>%
    dplyr::filter(Biodomain %in% top_bd_f)

  if (nrow(sub_ora_f) > 0) {
    plot_subdomain_distribution(
      data  = sub_ora_f,
      title = paste("Sub-BioDomain ORA — Females | Top", top_biodomains, "Domains")
    )
  } else {
    message("No sub-domain data available for females.")
  }
}
```

### 3.4 Sub-BioDomain ORA — Males

```{r ora_sub_male, fig.width=14, fig.height=16, eval=run_biodomain_ora}
if (!is.null(enr_ora_m) && nrow(enr_ora_m) > 0) {

  top_bd_m <- enr_ora_m %>%
    dplyr::filter(Biodomain != "none") %>%
    dplyr::count(Biodomain, sort = TRUE) %>%
    head(top_biodomains) %>%
    dplyr::pull(Biodomain)

  sub_ora_m <- enr_ora_m %>%
    dplyr::filter(Biodomain %in% top_bd_m)

  if (nrow(sub_ora_m) > 0) {
    plot_subdomain_distribution(
      data  = sub_ora_m,
      title = paste("Sub-BioDomain ORA — Males | Top", top_biodomains, "Domains")
    )
  } else {
    message("No sub-domain data available for males.")
  }
}
```

---

## 4. BioDomain — GSEA {.tabset .tabset-pills}

Gene Set Enrichment Analysis using all proteins (ranked by fold-change),
annotated with AD biological domains. Unlike ORA, GSEA uses the full ranked
list and does not require a significance cutoff for gene selection.

```{r gsea_check}
if (!run_biodomain_gsea) {
  cat("BioDomain GSEA skipped.\n",
      "Set run_biodomain_gsea <- TRUE and provide biodom_path and domain_colors_path to enable.\n")
}
```

### 4.1 GSEA — Females

```{r gsea_female, fig.width=14, fig.height=10, eval=run_biodomain_gsea}
enr_gsea_f <- enrichGSEA_BD_Protein(
  dat     = female_degs,
  bio_ref = bio_ref,
  lab_ref = lab_ref,
  org_db  = org_db_obj,
  seed    = gsea_seed
)

if (!is.null(enr_gsea_f) && nrow(enr_gsea_f) > 0) {
  plot_gsea_biodomain_distribution(enr_gsea_f, ran = nes_axis_limit)
} else {
  message("No GSEA results for females.")
}
```

### 4.2 GSEA — Males

```{r gsea_male, fig.width=14, fig.height=10, eval=run_biodomain_gsea}
enr_gsea_m <- enrichGSEA_BD_Protein(
  dat     = male_degs,
  bio_ref = bio_ref,
  lab_ref = lab_ref,
  org_db  = org_db_obj,
  seed    = gsea_seed
)

if (!is.null(enr_gsea_m) && nrow(enr_gsea_m) > 0) {
  plot_gsea_biodomain_distribution(enr_gsea_m, ran = nes_axis_limit)
} else {
  message("No GSEA results for males.")
}
```

### 4.3 Sub-BioDomain GSEA — Females

```{r gsea_sub_female, fig.width=14, fig.height=16, eval=run_biodomain_gsea}
if (!is.null(enr_gsea_f) && nrow(enr_gsea_f) > 0) {

  top_bd_gsea_f <- enr_gsea_f %>%
    dplyr::filter(Biodomain != "none", p.adjust < pval_cutoff) %>%
    dplyr::count(Biodomain, sort = TRUE) %>%
    head(top_biodomains) %>%
    dplyr::pull(Biodomain)

  sub_gsea_f <- enr_gsea_f %>%
    dplyr::filter(Biodomain %in% top_bd_gsea_f)

  if (nrow(sub_gsea_f) > 0) {
    plot_gsea_subdomain_distribution(sub_gsea_f, ran = nes_axis_limit)
  } else {
    message("No significant sub-domain GSEA results for females.")
  }
}
```

### 4.4 Sub-BioDomain GSEA — Males

```{r gsea_sub_male, fig.width=14, fig.height=16, eval=run_biodomain_gsea}
if (!is.null(enr_gsea_m) && nrow(enr_gsea_m) > 0) {

  top_bd_gsea_m <- enr_gsea_m %>%
    dplyr::filter(Biodomain != "none", p.adjust < pval_cutoff) %>%
    dplyr::count(Biodomain, sort = TRUE) %>%
    head(top_biodomains) %>%
    dplyr::pull(Biodomain)

  sub_gsea_m <- enr_gsea_m %>%
    dplyr::filter(Biodomain %in% top_bd_gsea_m)

  if (nrow(sub_gsea_m) > 0) {
    plot_gsea_subdomain_distribution(sub_gsea_m, ran = nes_axis_limit)
  } else {
    message("No significant sub-domain GSEA results for males.")
  }
}
```

---

## 5. Linear Modeling + Human Correlation {.tabset .tabset-pills}

Linear models test the effect of each genotype on protein abundance while
controlling for covariates (Age, Sex). This is complementary to the ANOVA
approach — ANOVA identifies pairwise differences, while LM estimates the
independent contribution of each genotype across the full dataset.

LM estimates are then correlated with human AD co-expression module signatures
and AD biological domain (BioDomain) reference data to assess translational
relevance.

```{r lm_check}
if (!run_lm_analysis) {
  cat("Linear modeling skipped.\n",
      "Set run_lm_analysis <- TRUE to enable.\n",
      "Also ensure run_module_corr and/or run_biodomain_ora is TRUE for human correlation.\n")
}
```

### 5.1 Run Linear Models

```{r lm_run, eval=run_lm_analysis}
# Load the abundance + traits data saved by notebook 01
processed_file <- file.path(data_dir, "Processed_DIA_Proteomics.RData")
if (!file.exists(processed_file))
  stop("Processed data not found at: ", processed_file,
       "\nPlease run 01_DIA_Proteomics_Pipeline.Rmd first.")

load(processed_file)   # loads: trait_data, clean_abundance, abundance_with_traits

# Set reference genotype level
abundance_with_traits <- abundance_with_traits %>%
  dplyr::mutate(
    Genotype = relevel(factor(Genotype), ref = lm_reference_genotype),
    Age      = as.numeric(Age),
    Sex      = factor(Sex)
  )

# Build model formula
formula_obj <- as.formula(paste("protein_col", lm_formula))

# Identify protein columns (all numeric columns after metadata)
meta_cols    <- c("animalName", "Genotype", "Sex", "Age", "Group")
protein_cols <- setdiff(names(abundance_with_traits), meta_cols)
protein_cols <- protein_cols[sapply(abundance_with_traits[protein_cols], is.numeric)]

cat("Running linear models for", length(protein_cols), "proteins...\n")

# Run LM for each protein
lm_summaries <- lapply(setNames(protein_cols, protein_cols), function(prot) {
  tryCatch({
    formula_str <- paste0("`", prot, "`", lm_formula)
    fit <- lm(as.formula(formula_str), data = abundance_with_traits)
    summary(fit)
  }, error = function(e) NULL)
})

# Remove failed models
lm_summaries <- Filter(Negate(is.null), lm_summaries)
cat("Successful models:", length(lm_summaries), "\n")

# Auto-detect coefficient names if not specified
if (is.null(lm_coef_names)) {
  lm_coef_names <- rownames(lm_summaries[[1]]$coefficients)
  cat("Auto-detected coefficients:\n")
  cat(paste0("  - ", lm_coef_names, collapse = "\n"), "\n")
}

# Extract estimates and p-values
lm_results <- extract_lm_stats(lm_summaries, coef_names = lm_coef_names)

cat("\nLM results preview:\n")
head(lm_results)
```

### 5.2 LM Results Summary

```{r lm_summary_table, eval=run_lm_analysis}
# Count significant variants per coefficient
lm_sig_summary <- lm_results %>%
  dplyr::group_by(Variant) %>%
  dplyr::summarise(
    Total_Proteins  = dplyr::n(),
    Significant     = sum(P.Value < pval_cutoff, na.rm = TRUE),
    Pct_Significant = round(100 * Significant / Total_Proteins, 1),
    Mean_Abs_Est    = round(mean(abs(Estimate), na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  dplyr::filter(!grepl("Intercept", Variant)) %>%
  dplyr::arrange(dplyr::desc(Significant))

lm_sig_summary %>%
  gt::gt() %>%
  gt::tab_header(
    title    = gt::md("**Linear Model Results Summary**"),
    subtitle = gt::md(paste0("Significance threshold: p < ", pval_cutoff))
  ) %>%
  gt::cols_label(
    Variant         = "Coefficient",
    Total_Proteins  = "Total Proteins",
    Significant     = "Significant (p<threshold)",
    Pct_Significant = "% Significant",
    Mean_Abs_Est    = "Mean |Estimate|"
  ) %>%
  gt::data_color(
    columns = Pct_Significant,
    method  = "numeric",
    palette = c("white", "#164B6E")
  ) %>%
  gt::tab_options(table.width = gt::pct(80),
                  column_labels.font.weight = "bold")
```

### 5.3 LM — Human Module Correlation

```{r lm_module_corr, fig.width=14, fig.height=7, eval=run_lm_analysis}
if (!run_module_corr) {
  message("LM module correlation skipped — set run_module_corr <- TRUE and provide module_data_file.")
} else {

  lm_module_cor <- correlate_lm_with_human_modules(
    lm_results      = lm_results,
    human_ref       = pmod.AD.Control,
    ordered_variant = order_variant_lm,
    module_order    = NULL
  ) 
  
  # Add group column from config if defined
  if (!is.null(lm_variant_groups) && !is.null(lm_module_cor)) {
    lm_module_cor <- lm_module_cor %>%
      dplyr::mutate(
        group = dplyr::recode(as.character(Variant), !!!lm_variant_groups,
                              .default = "Other")
      )

    if (!is.null(lm_group_order)) {
      lm_module_cor <- lm_module_cor %>%
        dplyr::mutate(group = factor(group, levels = lm_group_order))
    }
  }

  if (!is.null(lm_module_cor) && nrow(lm_module_cor) > 0) {
    protein_variant_corrplot_test(lm_module_cor)
  } else {
    message("No LM-module correlations computed (insufficient overlapping genes).")
  }
}
```

### 5.4 LM — BioDomain Correlation

```{r lm_biodomain_corr, fig.width=14, fig.height=7, eval=run_lm_analysis}
if (!run_biodomain_ora) {
  message("LM BioDomain correlation skipped — set run_biodomain_ora <- TRUE and provide biodom_path.")
} else {

  lm_biodom_cor <- corr_function_lm_biodomain(
    data      = lm_results,
    human_ref = bio_ref,
    sig       = pval_cutoff
  )

  if (!is.null(lm_biodom_cor) && nrow(lm_biodom_cor) > 0) {
    variant_corrplot_biodom(lm_biodom_cor)
  } else {
    message("No LM-BioDomain correlations computed (insufficient overlapping genes).")
  }
}
```

### 5.5 LM — Sub-BioDomain Correlation

```{r lm_subbiodomain_corr, fig.width=14, fig.height=16, eval=run_lm_analysis}
if (!run_biodomain_ora) {
  message("LM Sub-BioDomain skipped — set run_biodomain_ora <- TRUE.")
} else {

  lm_subbiodom_cor <- corr_function_lm_subbiodomain(
    lm_results = lm_results,
    human_ref  = bio_ref,
    sig        = pval_cutoff
  )

  if (!is.null(lm_subbiodom_cor) && nrow(lm_subbiodom_cor) > 0) {

    # Select top biodomains by number of significant sub-domains
    top_bd_lm <- lm_subbiodom_cor %>%
      dplyr::filter(significant) %>%
      dplyr::count(module, sort = TRUE) %>%
      head(top_biodomains) %>%
      dplyr::pull(module)

    lm_sub_plot <- lm_subbiodom_cor %>%
      dplyr::filter(module %in% top_bd_lm)

    if (nrow(lm_sub_plot) > 0) {
      variant_corrplot_subbiodom.Potrait(lm_sub_plot)
    } else {
      message("No significant sub-domain LM results to plot.")
    }
  } else {
    message("No LM Sub-BioDomain correlations computed.")
  }
}
```

### 5.6 LM — Scatter: Top Modules

```{r lm_module_scatter, fig.width=12, fig.height=8, eval=run_lm_analysis}
if (!run_module_corr) {
  message("LM module scatter skipped — set run_module_corr <- TRUE.")
} else if (exists("lm_module_cor") && !is.null(lm_module_cor) && nrow(lm_module_cor) > 0) {

  top_mods_lm <- lm_module_cor %>%
    dplyr::filter(significant) %>%
    dplyr::arrange(p_value) %>%
    dplyr::pull(Module) %>%
    unique() %>%
    head(top_modules)

  if (length(top_mods_lm) > 0) {
    plot_module_effect_scatter_correlations(
      deg_df       = lm_results,
      human_ref    = module_ref,
      mods_to_plot = top_mods_lm,
      sex_label    = "LM Estimate"
    )
  } else {
    message("No significant LM-module associations to plot.")
  }
}
```

---

## 6. Save Results

```{r save_results}
# Collect all results that were computed
results_to_save <- list()

if (run_module_corr) {
  results_to_save$module_cor_female <- cor_female
  results_to_save$module_cor_male   <- cor_male
}

if (run_biodomain_ora) {
  results_to_save$ora_female <- enr_ora_f
  results_to_save$ora_male   <- enr_ora_m
}

if (run_biodomain_gsea) {
  results_to_save$gsea_female <- enr_gsea_f
  results_to_save$gsea_male   <- enr_gsea_m
}

if (run_lm_analysis) {
  results_to_save$lm_results   <- lm_results
  results_to_save$lm_summaries <- lm_sig_summary
  if (exists("lm_module_cor"))    results_to_save$lm_module_cor    <- lm_module_cor
  if (exists("lm_biodom_cor"))    results_to_save$lm_biodom_cor    <- lm_biodom_cor
  if (exists("lm_subbiodom_cor")) results_to_save$lm_subbiodom_cor <- lm_subbiodom_cor
}

if (length(results_to_save) > 0) {
  save(results_to_save,
       file = file.path(result_dir, "Human_Translation_Results.Rdata"))
  message("Results saved to: ", file.path(result_dir, "Human_Translation_Results.Rdata"))
  cat("Objects saved:\n")
  cat(paste0("  - ", names(results_to_save), collapse = "\n"), "\n")
} else {
  message("No results to save — all analysis switches are set to FALSE.")
}
```

---

## 7. Session Info

```{r session_info, echo=TRUE}
sessionInfo()
```
