---
title: "Brain RNA-seq Analysis Pipeline"
author: "Ravi S Pandey"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, error = FALSE, echo = FALSE)
```

---

## ⚙️ USER CONFIGURATION — Edit This Section Before Running

```{r user_config, echo=TRUE}
# ── PATHS ─────────────────────────────────────────────────────────────────────
data_dir   <- "../data/"
result_dir <- "../results/"

# Input file names (place these in data_dir)
gene_counts_file    <- "rsem.merged.gene_counts.tsv"      # Raw gene count matrix
gene_tpm_file       <- "rsem.merged.gene_tpm.tsv"         # TPM matrix
isoform_tpm_file    <- "rsem.merged.transcript_tpm.tsv"   # Isoform TPM matrix
metadata_file       <- "metadata_validated.csv"            # Sample metadata CSV

# ── STUDY DESIGN ──────────────────────────────────────────────────────────────
# Control genotype name (must match values in metadata Genotype/line column)
control_genotype <- "Control_Genotype_Name"

# Case genotype names (must match values in metadata Genotype/line column)
case_genotypes <- c("Case_Genotype_1", "Case_Genotype_2", "Case_Genotype_3")

# Age groups in study (numeric, as they appear in metadata)
age_groups <- c(4, 12)

# Sex levels
sex_levels <- c("Female", "Male")

# ── METADATA COLUMN MAPPING ───────────────────────────────────────────────────
# Column names in your metadata CSV — adjust to match your file
col_sample_id  <- "climbID"      # Unique sample identifier
col_sex        <- "sex_climb"    # Sex column
col_age        <- "age_m"        # Age column (numeric months)
col_genotype   <- "line"         # Genotype/line column
col_name       <- "sampleName"   # Optional sample name column

# Age threshold for dichotomizing to timepoints (e.g., >10 months → 12M)
age_threshold     <- 10
age_above_thresh  <- 12
age_below_thresh  <- 4

# Sample IDs to exclude (e.g., failed QC, wrong genotype) — set to NULL if none
samples_to_exclude <- NULL    # e.g., c(424862, 424866, 424858)

# ── TRANSGENE HANDLING ────────────────────────────────────────────────────────
# Set TRUE if your model has human transgenes that need to be merged with
# their mouse orthologs before analysis (e.g., human MAPT + mouse Mapt)
has_transgenes <- FALSE

# If has_transgenes = TRUE, define the human→mouse Ensembl ID merge pairs.
# Each entry: human_id will be summed into mouse_id, then human_id removed.
# Example for MAPT: human ENSG00000186868 merged into mouse ENSMUSG00000018411
transgene_merge_pairs <- list(
  # list(human_id = "ENSG00000186868", mouse_id = "ENSMUSG00000018411")
)

# Ensembl IDs to drop entirely after merging (e.g., non-coding human transgenes)
# Set to NULL if none
transgene_ids_to_drop <- NULL   # e.g., "ENSG0"  (prefix match via %like%)

# Gene IDs for transgene visualization (raw counts and TPM boxplots)
transgene_viz_ids <- c()   # e.g., c("ENSG00000186868", "ENSMUSG00000018411")

# ── ISOFORM ANALYSIS ──────────────────────────────────────────────────────────
# Set TRUE only if you have isoform TPM data and want to visualize isoforms
run_isoform_analysis <- FALSE

# Human gene Ensembl ID whose isoforms to visualize (e.g., MAPT = ENSG00000186868)
isoform_gene_id <- ""

# Lookup table: Transcript ID → readable label
# Example: c("ENST00000334239" = "ENST00000334239 (0N3R)", ...)
isoform_lookup <- c()

# ── SEX VALIDATION GENES ──────────────────────────────────────────────────────
# Ensembl IDs for female-specific and male-specific genes to validate sex assignment
sex_validation_ids <- c(
  "ENSMUSG00000086503",   # Xist (female)
  "ENSMUSG00000069045"    # Ddx3y (male)
)
sex_validation_labels <- c(
  "ENSMUSG00000086503" = "Mouse Xist",
  "ENSMUSG00000069045" = "Mouse Ddx3y"
)

# ── PRE-FILTERING THRESHOLDS ──────────────────────────────────────────────────
tpm_min_count     <- 0.1    # Minimum TPM to count a gene as expressed
tpm_min_samples   <- 5      # Minimum number of samples above tpm_min_count

# ── SPECIES ───────────────────────────────────────────────────────────────────
# Organism database for gene symbol mapping
# Mouse: "org.Mm.eg.db" | Human: "org.Hs.eg.db" | Rat: "org.Rn.eg.db"
organism_db  <- "org.Mm.eg.db"
kegg_organism <- "mmu"        # "hsa" for human, "rno" for rat

# ── DESEQ2 SETTINGS ───────────────────────────────────────────────────────────
deseq2_alpha      <- 0.05    # Significance threshold for DESeq2 results()
deseq2_min_count  <- 10      # Minimum row sum to retain a gene (pre-filter)
deseq2_lfc_cutoff <- 0       # Log2FC threshold for DEG barplots/volcano

# ── PCA SETTINGS ──────────────────────────────────────────────────────────────
pca_ntop <- 1000    # Number of top variable genes for PCA
pca_remove_xist <- TRUE  # Set TRUE to remove Xist (ENSMUSG00000086503) before PCA

# ── KEGG SETTINGS ─────────────────────────────────────────────────────────────
kegg_top_n    <- 15     # Number of top pathways to display
kegg_pval_cutoff <- 0.05
```

```{r load_organism_db, echo=TRUE}
# Load organism db — done once
library(organism_db, character.only = TRUE)
org_db_obj <- get(organism_db)
```

```{r source_helpers}
source("../R/Helper_Functions_RNA.R")
```

---

# {.tabset .tabset-fade}

## 1. Data Inspection

### 1.1 Study Overview

> Data can be downloaded from the project Synapse page. Update paths in the
> USER CONFIGURATION block before running.

### 1.2 Load & Format Metadata

```{r metadata, echo=TRUE}
meta_raw <- read.csv(file.path(data_dir, metadata_file)) %>%
  dplyr::distinct(.data[[col_sample_id]], .keep_all = TRUE)

# Exclude failed QC samples if specified
if (!is.null(samples_to_exclude)) {
  meta_raw <- meta_raw %>%
    dplyr::filter(!(.data[[col_sample_id]] %in% samples_to_exclude))
}

# Standardize column names and dichotomize age
meta <- meta_raw %>%
  dplyr::mutate(
    Age = ifelse(.data[[col_age]] > age_threshold, age_above_thresh, age_below_thresh)
  ) %>%
  dplyr::select(
    Names    = !!col_sample_id,
    sampleName = dplyr::any_of(col_name),
    Sex      = !!col_sex,
    Age,
    Genotype = !!col_genotype
  ) %>%
  dplyr::filter(Genotype %in% c(control_genotype, case_genotypes))

# Study population summary
dplyr::count(meta, Genotype, Age, Sex) %>%
  tidyr::pivot_wider(names_from = Sex, values_from = n, values_fill = 0) %>%
  gt::gt() %>%
  gt::tab_header(title = gt::md("**Study Population**"))
```

### 1.3 Load Count Matrices

```{r load_counts, echo=TRUE}
gene_counts_raw <- readr::read_tsv(
  file.path(data_dir, gene_counts_file), show_col_types = FALSE
) %>%
  dplyr::select(-dplyr::any_of("transcript_id(s)"))

gene_tpm_raw <- readr::read_tsv(
  file.path(data_dir, gene_tpm_file), show_col_types = FALSE
) %>%
  dplyr::select(-dplyr::any_of("transcript_id(s)"))

cat("Raw count matrix:", nrow(gene_counts_raw), "genes x",
    ncol(gene_counts_raw) - 1, "samples\n")
```

### 1.4 Transgene Inspection & Merging

```{r transgene_inspect, eval=has_transgenes, fig.width=11, fig.height=10}
# Plot raw counts for transgene genes of interest
if (length(transgene_viz_ids) > 0) {
  tg_plot <- gene_counts_raw %>%
    dplyr::filter(gene_id %in% transgene_viz_ids) %>%
    tidyr::pivot_longer(cols = -gene_id, names_to = "Names", values_to = "counts") %>%
    dplyr::mutate(Names = as.integer(Names)) %>%
    dplyr::left_join(meta, by = "Names")

  ggplot2::ggplot(tg_plot, ggplot2::aes(x = Genotype, y = counts, fill = factor(Age))) +
    ggplot2::geom_boxplot(ggplot2::aes(color = factor(Age)), alpha = 0.2,
                          outlier.shape = NA) +
    ggplot2::geom_point(ggplot2::aes(shape = Sex, color = factor(Age)),
                        position = ggplot2::position_jitterdodge(
                          jitter.width = 0.2, dodge.width = 0.75),
                        alpha = 0.7, size = 2) +
    ggplot2::facet_wrap(~gene_id, scales = "free_y") +
    ggplot2::labs(x = "Genotype", y = "Raw Counts",
                  title = "Transgene Expression — Raw Counts") +
    ggplot2::theme_minimal() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
}
```

```{r transgene_merge, echo=TRUE}
# Remove excluded samples from count matrix
exclude_cols <- as.character(samples_to_exclude)

gene_counts <- gene_counts_raw %>%
  dplyr::select(-dplyr::any_of(exclude_cols)) %>%
  tibble::column_to_rownames("gene_id")

tpm_mat <- gene_tpm_raw %>%
  dplyr::select(-dplyr::any_of(exclude_cols)) %>%
  tibble::column_to_rownames("gene_id")

# Merge human transgene counts into mouse orthologs
if (has_transgenes && length(transgene_merge_pairs) > 0) {
  for (pair in transgene_merge_pairs) {
    gene_counts[pair$mouse_id, ] <-
      gene_counts[pair$mouse_id, ] + gene_counts[pair$human_id, ]
    gene_counts <- gene_counts[!rownames(gene_counts) %in% pair$human_id, ]

    tpm_mat[pair$mouse_id, ] <-
      tpm_mat[pair$mouse_id, ] + tpm_mat[pair$human_id, ]
    tpm_mat <- tpm_mat[!rownames(tpm_mat) %in% pair$human_id, ]
  }
  cat("Transgene merging complete.\n")
}

# Drop any remaining human/unwanted genes by prefix
if (!is.null(transgene_ids_to_drop)) {
  keep_rows <- !grepl(paste0("^", transgene_ids_to_drop), rownames(gene_counts))
  gene_counts <- gene_counts[keep_rows, ]
  tpm_mat     <- tpm_mat[keep_rows, ]
  cat("Dropped genes matching prefix:", transgene_ids_to_drop, "\n")
}

cat("Final count matrix:", nrow(gene_counts), "genes x", ncol(gene_counts), "samples\n")
```

### 1.5 Sex Validation

```{r sex_validation, fig.width=9, fig.height=4}
if (length(sex_validation_ids) > 0) {
  sex_val <- tpm_mat %>%
    tibble::rownames_to_column("gene_id") %>%
    dplyr::filter(gene_id %in% sex_validation_ids) %>%
    tidyr::pivot_longer(cols = -gene_id, names_to = "Names", values_to = "tpm") %>%
    dplyr::mutate(Names = as.integer(Names)) %>%
    dplyr::left_join(meta, by = "Names") %>%
    dplyr::mutate(
      Age    = factor(Age, levels = age_groups),
      symbol = dplyr::recode(gene_id, !!!sex_validation_labels)
    )

  ggplot2::ggplot(sex_val, ggplot2::aes(x = Sex, y = tpm,
                                         color = factor(Age), shape = Genotype)) +
    ggplot2::geom_boxplot() +
    ggplot2::geom_point(position = ggplot2::position_jitterdodge()) +
    ggplot2::facet_wrap(~symbol, scales = "free") +
    ggplot2::theme_bw() +
    ggplot2::labs(title = "Sex Validation — Xist (F) and Ddx3y (M)", color = "Age")
}
```

### 1.6 Zero Count Filtering Summary

```{r zero_filter}
gene_sums <- data.frame(
  gene_id = rownames(gene_counts),
  sums    = Matrix::rowSums(gene_counts)
)
cat("Genes with zero counts across all samples:",
    sum(gene_sums$sums == 0), "\n")
cat("These will be removed by DESeq2 pre-filtering.\n")
```

### 1.7 Pre-filter & Save Processed Data

```{r prefilter, eval=FALSE, echo=TRUE}
# TPM-based pre-filtering and log2 transformation
TPMcount.filter <- tpm_mat[rowSums(tpm_mat > tpm_min_count) > tpm_min_samples, ]
rawcountdata    <- gene_counts[, as.character(meta$Names)]
mydat_TPM       <- as.data.frame(log2(
  TPMcount.filter[, as.character(meta$Names)] + 1
))

# Format metadata for DESeq2
metadata_rna <- as.data.frame(meta)
rownames(metadata_rna) <- metadata_rna$Names

save(metadata_rna, rawcountdata, mydat_TPM,
     file = file.path(data_dir, "ProcessedData_RNAseq.RData"))
cat("Processed data saved.\n")
```

---

## 2. Isoform Expression

```{r isoform_check}
if (!run_isoform_analysis) {
  cat("Isoform analysis skipped.\n",
      "Set run_isoform_analysis <- TRUE and provide isoform_gene_id and isoform_lookup to enable.\n")
}
```

```{r isoform_load, eval=run_isoform_analysis}
load(file.path(data_dir, "ProcessedData_RNAseq.RData"))

tpm_isoform <- read.delim2(
  file.path(data_dir, isoform_tpm_file), check.names = FALSE
)

isodata_meta <- tpm_isoform %>%
  dplyr::filter(gene_id == isoform_gene_id) %>%
  dplyr::select(-gene_id) %>%
  tidyr::pivot_longer(cols = -transcript_id, names_to = "Names", values_to = "TPM") %>%
  dplyr::mutate(
    Names  = as.numeric(Names),
    TPM    = as.numeric(TPM),
    logTPM = log2(TPM + 1),
    logTPM = ifelse(is.infinite(logTPM) | is.na(logTPM), 0, logTPM)
  ) %>%
  dplyr::inner_join(
    metadata_rna %>%
      dplyr::mutate(Group = paste0(Genotype, "_", Sex, "_", Age, "M")),
    by = "Names"
  )
```

### 2.1 Selected Isoforms

```{r isoform_plot_selected, fig.width=18, fig.height=12, eval=run_isoform_analysis}
if (length(isoform_lookup) > 0) {
  df_plot <- isodata_meta %>%
    dplyr::filter(transcript_id %in% names(isoform_lookup)) %>%
    dplyr::mutate(
      transcript_label = factor(isoform_lookup[transcript_id],
                                levels = isoform_lookup)
    )

  # Calculate group means for text labels
  means_df <- df_plot %>%
    dplyr::group_by(Group, transcript_label) %>%
    dplyr::summarise(mean_val = round(mean(logTPM), 2), .groups = "drop")

  ggplot2::ggplot(df_plot, ggplot2::aes(x = Group, y = logTPM, fill = Group)) +
    ggplot2::geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 1/3) +
    ggplot2::stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1),
                          geom = "errorbar", width = 0.2) +
    ggplot2::stat_summary(fun = mean, geom = "point", size = 2) +
    ggplot2::geom_text(data = means_df,
                       ggplot2::aes(y = mean_val + 0.5, label = mean_val),
                       size = 3, hjust = 0.5) +
    ggplot2::facet_wrap(~transcript_label, ncol = 3) +
    ggplot2::labs(title = paste("Isoforms of", isoform_gene_id),
                  y = "Expression (log2 TPM)", x = "") +
    ggplot2::theme_bw(base_size = 14) +
    ggplot2::theme(
      plot.title   = ggplot2::element_text(hjust = 0.5, face = "bold"),
      axis.text.x  = ggplot2::element_text(angle = 90, hjust = 1, face = "bold"),
      strip.text   = ggplot2::element_text(size = 12, face = "bold"),
      legend.position = "none"
    )
}
```

---

## 3. Principal Component Analysis

```{r load_for_pca}
load(file.path(data_dir, "ProcessedData_RNAseq.RData"))
```

### 3.1 DESeq2 VST PCA

```{r pca_deseq2, fig.width=10, fig.height=8}
dds_pca <- DESeq2::DESeqDataSetFromMatrix(
  countData = as.matrix(rawcountdata),
  colData   = metadata_rna,
  design    = as.formula(paste0("~ Genotype"))
)
dds_pca <- dds_pca[rowSums(DESeq2::counts(dds_pca) > 1) >= 6, ]
dds_pca <- DESeq2::DESeq(dds_pca, parallel = TRUE)
vsd     <- DESeq2::varianceStabilizingTransformation(dds_pca, blind = FALSE)
vsd_mat <- DESeq2::assay(vsd)

# Optionally remove Xist before PCA (avoids sex dominating PC1)
if (pca_remove_xist) {
  vsd_mat <- vsd_mat[!rownames(vsd_mat) %in% "ENSMUSG00000086503", ]
}

pca_df      <- plotPCA.df(vsd_mat, metadata = metadata_rna,
                           intgroup = c("Names", "Genotype", "Age", "Sex"),
                           returnData = TRUE, ntop = pca_ntop)
percent_var <- round(100 * attr(pca_df, "percentVar"))

# Plot 1: Genotype + Sex
ggplot2::ggplot(pca_df, ggplot2::aes(x = PC1, y = PC2,
                                      color = Genotype, shape = Sex)) +
  ggplot2::geom_point(size = 4, alpha = 0.8) +
  ggplot2::geom_text(ggplot2::aes(label = Age), hjust = 0.5, vjust = 2.5,
                     size = 3.5, show.legend = FALSE) +
  ggplot2::labs(
    x = paste0("PC1: ", percent_var[1], "% variance"),
    y = paste0("PC2: ", percent_var[2], "% variance"),
    title = "PCA — Variance Stabilized Counts",
    subtitle = paste0("Top ", pca_ntop, " variable genes")
  ) +
  ggplot2::theme_bw(base_size = 16) +
  ggplot2::theme(
    plot.title    = ggplot2::element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = ggplot2::element_text(hjust = 0.5),
    axis.text     = ggplot2::element_text(face = "bold", color = "black"),
    legend.position = "right",
    panel.grid.minor = ggplot2::element_blank()
  ) +
  ggplot2::scale_color_brewer(palette = "Set1")
```

```{r pca_age_sex, fig.width=10, fig.height=8}
# Plot 2: Age + Genotype
ggplot2::ggplot(pca_df, ggplot2::aes(x = PC1, y = PC2,
                                      color = factor(Age), shape = Genotype)) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_text(ggplot2::aes(label = Names), hjust = 0.5, vjust = 2,
                     size = 3.5) +
  ggplot2::labs(
    x = paste0("PC1: ", percent_var[1], "% variance"),
    y = paste0("PC2: ", percent_var[2], "% variance"),
    color = "Age (months)"
  ) +
  ggplot2::theme_bw(base_size = 16) +
  ggplot2::theme(panel.grid.minor = ggplot2::element_blank())
```

### 3.2 PCAtools Biplot + Eigencor

```{r pca_tools, fig.width=9, fig.height=6}
key <- metadata_rna[colnames(mydat_TPM), ]
p   <- PCAtools::pca(mydat_TPM, metadata = key, removeVar = 0.1)

PCAtools::biplot(p, colby = "Genotype", shape = "Sex",
                 legendPosition = "right", title = "PCA biplot — TPM",
                 lab = p$metadata$Age, drawConnectors = FALSE)

PCAtools::eigencorplot(p, metavars = c("Genotype", "Age", "Sex"))
```

### 3.3 Final Study Population

```{r final_population}
dplyr::count(metadata_rna, Genotype, Age, Sex) %>%
  tidyr::pivot_wider(names_from = Sex, values_from = n, values_fill = 0) %>%
  gt::gt() %>%
  gt::tab_header(title = gt::md("**Final Study Population**"))
```

---

## 4. Differential Expression Analysis

### 4.1 Setup Comparisons

```{r deseq2_setup, echo=TRUE}
load(file.path(data_dir, "ProcessedData_RNAseq.RData"))

metadata <- metadata_rna %>%
  dplyr::mutate(
    Genotype = factor(Genotype,
                      levels = c(control_genotype, case_genotypes)),
    Group = paste0(Genotype, "_", Sex, "_", Age, "M")
  )

# Auto-build valid comparisons (only pairs present in data)
available_groups <- unique(metadata$Group)

comp_table <- expand.grid(
  case_genotype = case_genotypes,
  sex           = sex_levels,
  age           = age_groups,
  stringsAsFactors = FALSE
) %>%
  dplyr::mutate(
    control = paste0(control_genotype, "_", sex, "_", age, "M"),
    case    = paste0(case_genotype,    "_", sex, "_", age, "M")
  ) %>%
  dplyr::filter(
    control %in% available_groups,
    case    %in% available_groups
  ) %>%
  dplyr::select(control, case)

cat("Valid comparisons:", nrow(comp_table), "\n")
print(comp_table)
```

### 4.2 Run DESeq2

```{r run_deseq2, eval=FALSE, echo=TRUE}
res_deseq <- run_experiment_analysis(
  counts          = rawcountdata,
  meta            = metadata,
  comp_table      = comp_table,
  target_col      = "Genotype",
  experiment_name = "RNAseq"
)

save(res_deseq,
     file = file.path(result_dir, "DESeq2_Results_RNAseq.RData"))
cat("DESeq2 results saved.\n")
```

```{r load_deseq2}
load(file.path(result_dir, "DESeq2_Results_RNAseq.RData"))
cat("Models loaded:", length(res_deseq$list), "\n")
```

### 4.3 DEG Summary Table

```{r deg_summary}
degs_up   <- purrr::map(res_deseq$list,
                         ~length(which(.x$padj < deseq2_alpha &
                                       .x$log2FoldChange > deseq2_lfc_cutoff)))
degs_down <- purrr::map(res_deseq$list,
                         ~length(which(.x$padj < deseq2_alpha &
                                       .x$log2FoldChange < -deseq2_lfc_cutoff)))

deg_summary <- data.frame(
  comparison = names(degs_up),
  Up         = as.integer(unlist(degs_up)),
  Down       = as.integer(unlist(degs_down))
) %>%
  dplyr::mutate(Sex = ifelse(grepl("Female", comparison), "Female", "Male"))

deg_summary %>%
  gt::gt() %>%
  gt::tab_header(
    title    = gt::md("**DEG Summary**"),
    subtitle = gt::md(paste0("FDR < ", deseq2_alpha,
                              " | |Log2FC| > ", deseq2_lfc_cutoff))
  ) %>%
  gt::tab_spanner(label = "Count", columns = c(Up, Down))
```

### 4.4 DEG Barplot

```{r deg_barplot, fig.width=8, fig.height=6}
my_colors <- c("Up_DEGs" = "#D55E00", "Down_DEGs" = "#0072B2")

g1 <- deg_summary %>%
  dplyr::rename(Up_DEGs = Up, Down_DEGs = Down) %>%
  tidyr::pivot_longer(cols = c(Up_DEGs, Down_DEGs),
                      names_to = "Group", values_to = "DEGs") %>%
  dplyr::mutate(
    comparison = stringr::str_remove_all(comparison, "Female_|Male_")
  )

barplot_fn(g1)
```

### 4.5 Volcano Plots

```{r volcano, fig.width=8, fig.height=8}
for (i in names(res_deseq$list)) {
  df <- as.data.frame(res_deseq$list[[i]])

  print(
    EnhancedVolcano::EnhancedVolcano(df,
      x = "log2FoldChange", y = "padj",
      lab = df$symbol, title = i, subtitle = "",
      FCcutoff  = deseq2_lfc_cutoff,
      pCutoff   = deseq2_alpha,
      xlim = c(min(df$log2FoldChange, na.rm = TRUE) - 1,
               max(df$log2FoldChange, na.rm = TRUE) + 1),
      ylim = c(0, max(-log10(df$padj), na.rm = TRUE) + 2)
    )
  )
}
```

---

## 5. Functional Analysis — KEGG

Over-representation analysis of DEGs against KEGG pathways using
`clusterProfiler::compareCluster()`.

```{r kegg, fig.height=9, fig.width=11}
res_all    <- res_deseq$df
sig_genes  <- dplyr::filter(res_all, padj < deseq2_alpha)
sig_groups <- unique(sig_genes$group)

if (length(sig_groups) > 0) {
  kegg_enrichment_analysis(
    groups      = sig_groups,
    deg_results = sig_genes,
    top_n       = kegg_top_n
  )
} else {
  message("No significant DEGs found for KEGG enrichment.")
}
```

---

## 6. Save Processed Results

```{r save_results}
save(res_deseq, deg_summary,
     file = file.path(result_dir, "DESeq2_Results_RNAseq.RData"))
cat("All results saved to:", file.path(result_dir, "DESeq2_Results_RNAseq.RData"), "\n")
```

---

## Session Info

```{r session_info, echo=TRUE}
sessionInfo()
```
