---
title: "RNA-seq Human Translation Analysis"
author: "Ravi S Pandey"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, error = FALSE, echo = FALSE)
```

> **Prerequisite:** Run `03_RNAseq_Pipeline.Rmd` first to generate
> `DESeq2_Results_RNAseq.RData` and `ProcessedData_RNAseq.RData`.

---

## ⚙️ USER CONFIGURATION — Edit This Section Before Running

```{r user_config, echo=TRUE}
# ── PATHS ─────────────────────────────────────────────────────────────────────
data_dir   <- "../data/"
result_dir <- "../results/"

# Input files from notebook 03
deseq2_file    <- "DESeq2_Results_RNAseq.RData"
processed_file <- "ProcessedData_RNAseq.RData"

# ── HUMAN REFERENCE DATA ──────────────────────────────────────────────────────
# AMP-AD module data — combined (used for logFC correlation)
# Expected to contain: ampad_modules_fc, module_clusters, mod
ampad_module_file <- ""        # e.g., "data/AMPADModuleData.RData"

# AMP-AD module data — sex-stratified (female/male separate)
# Expected to contain: ampad_modules_fc.female, ampad_modules_fc.male
ampad_sexwise_file <- ""       # e.g., "data/AMPAD_module_Sexwise.RData"

# AD subtype data — Nikhil et al. (ROSMAP, Mayo, MSBB)
# Expected to contain: ampad_subtype_fc, module_cohort, subtypes
nikhil_subtype_file <- ""      # e.g., "data/Nikhil_SubtypeData.RData"

# AD subtype data — Neff et al. (MSBB)
# Expected to contain: neff_subtype_fc
neff_subtype_file <- ""        # e.g., "data/Neff_Subtype.RData"

# BioDomain reference files
# biodom_path: mouse biodomain RDS (annotated_biodomains_mmus.rds)
# biodom_rna: used for ORA/GSEA (mouse-specific annotations)
biodom_path        <- ""       # e.g., "data/annotated_biodomains_mmus.rds"
domain_colors_path <- ""       # e.g., "data/domain_colors.csv"

# Human TREAT-AD omics scores (optional — for TE-based correlation)
treat_ad_file <- ""            # e.g., "data/OmicsScores_TREATAD.csv"

# ── ANALYSIS SWITCHES ─────────────────────────────────────────────────────────
run_module_corr_fc    <- FALSE   # AMP-AD module correlation using DESeq2 logFC
run_module_corr_lm    <- FALSE   # AMP-AD module correlation using LM estimates
run_subtype_corr      <- FALSE   # AD subtype correlation (Nikhil + Neff)
run_lm_analysis       <- FALSE   # Linear modeling on TPM data
run_biodomain_ora     <- FALSE   # GO-ORA with BioDomain annotation
run_biodomain_gsea    <- FALSE   # GO-GSEA with BioDomain annotation

# ── STATISTICAL THRESHOLDS ────────────────────────────────────────────────────
pval_cutoff   <- 0.05
lfc_cutoff    <- 0
nes_axis_limit <- 3

# ── LINEAR MODEL CONFIGURATION ────────────────────────────────────────────────
# Formula for LM: gene expression ~ covariates
# Genotype dummy variables are created from case_genotypes below
lm_formula <- "~ sex + Age + {case_dummies}"   # placeholder; built programmatically

# Genotype names as they will appear in LM coefficients
# These must match how R names them after dummy coding (e.g., paste0("Genotype", case))
case_genotypes <- c("Case_Genotype_1", "Case_Genotype_2", "Case_Genotype_3")
control_genotype <- "Control_Genotype_Name"

# Age subsets for stratified LM (NULL = all samples together)
lm_age_subsets <- list(
  all  = NULL,   # NULL means use all samples
  "4M" = 4,
  "12M" = 12
)

# Readable labels for LM coefficients — used in plots
# Keys must exactly match the coefficient names produced by your LM
lm_term_labels <- c(
  "(Intercept)" = "Intercept",
  "sex"         = "Sex (Male)",
  "Age12"       = "Age_12M"
  # Add genotype terms here, e.g.:
  # "IVS16"     = "Case_Genotype_1"
)

# Display order for variants in plots
order_variant_lm <- NULL   # e.g., c("Sex (Male)", "Age_12M", "Case_Genotype_1")

# Group mapping for faceting LM plots (Variant → group label)
lm_variant_groups <- list(
  "Sex (Male)" = "Covariate",
  "Age_12M"    = "Covariate"
  # Genotype terms → "Genotype Effects"
)
lm_group_order <- c("Genotype Effects", "Covariate")

# ── PLOT OPTIONS ──────────────────────────────────────────────────────────────
order_model_female <- NULL   # Factor order for female models in corrplots
order_model_male   <- NULL   # Factor order for male models in corrplots
top_biodomains     <- 5      # Top N biodomains for sub-domain plots
```

```{r source_helpers}
source("../R/Helper_Functions_RNA.R")
```

---

# {.tabset .tabset-fade}

## 1. Load Data

### 1.1 DESeq2 Results from Notebook 03

```{r load_deseq2}
deseq2_path <- file.path(result_dir, deseq2_file)
if (!file.exists(deseq2_path))
  stop("DESeq2 results not found at: ", deseq2_path,
       "\nPlease run 03_RNAseq_Pipeline.Rmd first.")

load(deseq2_path)
cat("Comparisons loaded:", length(res_deseq$list), "\n")
cat("Total DEG rows:", nrow(res_deseq$df), "\n")
```

### 1.2 Processed TPM Data from Notebook 03

```{r load_processed}
proc_path <- file.path(data_dir, processed_file)
if (!file.exists(proc_path))
  stop("Processed data not found at: ", proc_path,
       "\nPlease run 03_RNAseq_Pipeline.Rmd first.")

load(proc_path)   # loads: metadata_rna, rawcountdata, mydat_TPM
cat("Samples in TPM matrix:", ncol(mydat_TPM), "\n")
cat("Genes in TPM matrix:", nrow(mydat_TPM), "\n")
```

### 1.3 Load Human Reference Data

```{r load_human_refs}
# Load only the files that are needed by enabled analyses
if (run_module_corr_fc || run_module_corr_lm) {
  if (file.exists(ampad_module_file))  load(ampad_module_file)
  if (file.exists(ampad_sexwise_file)) load(ampad_sexwise_file)
}

if (run_subtype_corr) {
  if (file.exists(nikhil_subtype_file)) load(nikhil_subtype_file)
  if (file.exists(neff_subtype_file))   load(neff_subtype_file)
}

if (run_biodomain_ora || run_biodomain_gsea) {
  if (file.exists(biodom_path))
    biodom_mouse <- readRDS(biodom_path)
  if (file.exists(domain_colors_path))
    dom.lab <- readr::read_csv(domain_colors_path, show_col_types = FALSE)
}
```

### 1.4 Define Model Order

```{r model_order}
# Define display order for all models — update to match your comparisons
all_models <- names(res_deseq$list)

if (is.null(order_model_female))
  order_model_female <- all_models[grepl("Female", all_models)]
if (is.null(order_model_male))
  order_model_male   <- all_models[grepl("Male", all_models)]

cat("Female models:", length(order_model_female), "\n")
cat("Male models:  ", length(order_model_male), "\n")
```

---

## 2. AMP-AD Module Correlation — logFC {.tabset .tabset-pills}

Pearson correlation between mouse DESeq2 log2 fold-changes and human AMP-AD
co-expression module signatures. Black squares indicate p < 0.05.

```{r module_fc_check}
if (!run_module_corr_fc) {
  cat("Module logFC correlation skipped.\n",
      "Set run_module_corr_fc <- TRUE and provide ampad_module_file.\n")
}
```

### 2.1 Female — Module logFC

```{r corr_female_fc, fig.width=24, fig.height=6, eval=run_module_corr_fc}
corr_res_female <- correlation_analysis(
  mouse_res   = res_deseq$df %>% dplyr::filter(group %in% order_model_female),
  human_data  = ampad_modules_fc.female,
  module_meta = module_clusters,
  mod_order   = mod,
  sex_order   = order_model_female
)

if (!is.null(corr_res_female) && nrow(corr_res_female) > 0) {
  magora_corrplot_colored(corr_res_female)
} else {
  message("No female module correlations computed.")
}
```

### 2.2 Male — Module logFC

```{r corr_male_fc, fig.width=24, fig.height=6, eval=run_module_corr_fc}
corr_res_male <- correlation_analysis(
  mouse_res   = res_deseq$df %>% dplyr::filter(group %in% order_model_male),
  human_data  = ampad_modules_fc.male,
  module_meta = module_clusters,
  mod_order   = mod,
  sex_order   = order_model_male
)

if (!is.null(corr_res_male) && nrow(corr_res_male) > 0) {
  magora_corrplot_colored(corr_res_male)
} else {
  message("No male module correlations computed.")
}
```

---

## 3. Linear Modeling {.tabset .tabset-pills}

Per-gene linear models on log2 TPM data, estimating the independent effect
of each genotype variant while controlling for sex and age.

```{r lm_check}
if (!run_lm_analysis) {
  cat("Linear modeling skipped.\n",
      "Set run_lm_analysis <- TRUE to enable.\n")
}
```

### 3.1 Run Linear Models

```{r lm_run, eval=run_lm_analysis, echo=TRUE}
metadata <- metadata_rna %>%
  dplyr::mutate(
    Genotype = factor(Genotype, levels = c(control_genotype, case_genotypes)),
    Age      = factor(Age),
    Names    = as.character(Names)
  )

# Build long-format data with dummy coding
mydat_long <- mydat_TPM %>%
  as.data.frame() %>%
  dplyr::mutate(Gene = map_gene_ids(., "ENSEMBL", "SYMBOL")) %>%
  tidyr::pivot_longer(cols = -Gene, names_to = "Names", values_to = "value") %>%
  dplyr::filter(!is.na(Gene)) %>%
  dplyr::left_join(metadata, by = "Names") %>%
  dplyr::mutate(sex = as.numeric(Sex == "Male"))

# Add dummy variables for each case genotype
for (g in case_genotypes) {
  col_name <- make.names(g)
  mydat_long[[col_name]] <- as.numeric(mydat_long$Genotype == g)
}

# Build LM formula dynamically
case_dummy_cols <- make.names(case_genotypes)
lm_formula_str  <- paste0("value ~ sex + Age + ",
                           paste(case_dummy_cols, collapse = " + "))
cat("LM formula:", lm_formula_str, "\n")

# Run LM for all samples
run_lm_subset <- function(data, formula_str) {
  data %>%
    dplyr::group_by(Gene) %>%
    tidyr::nest() %>%
    dplyr::mutate(
      model  = purrr::map(data, ~lm(as.formula(formula_str), data = .x)),
      tidied = purrr::map(model, broom::tidy)
    ) %>%
    tidyr::unnest(tidied) %>%
    dplyr::ungroup() %>%
    dplyr::select(-data, -model) %>%
    dplyr::mutate(term = dplyr::recode(term, !!!lm_term_labels))
}

lm_results_all <- run_lm_subset(mydat_long, lm_formula_str)

# Run for each age subset
lm_results_by_age <- list()
for (age_label in names(lm_age_subsets)) {
  age_val <- lm_age_subsets[[age_label]]
  if (is.null(age_val)) next

  sub_formula <- paste0("value ~ sex + ", paste(case_dummy_cols, collapse = " + "))
  lm_results_by_age[[age_label]] <- run_lm_subset(
    mydat_long %>% dplyr::filter(Age == age_val),
    sub_formula
  )
  cat("LM completed for age:", age_label, "\n")
}

# Format for downstream functions
format_lm_effects <- function(tidy_df) {
  tidy_df %>%
    dplyr::filter(term != "Intercept") %>%
    dplyr::select(Gene, Variant = term, value = estimate, p.value)
}

rna_effects_all <- format_lm_effects(lm_results_all)
rna_effects_by_age <- purrr::map(lm_results_by_age, format_lm_effects)

save(rna_effects_all, rna_effects_by_age,
     file = file.path(result_dir, "LM_Results_RNAseq.RData"))
cat("LM results saved.\n")
```

```{r load_lm_results, eval=run_lm_analysis}
load(file.path(result_dir, "LM_Results_RNAseq.RData"))
```

### 3.2 LM Summary

```{r lm_summary, eval=run_lm_analysis}
lm_sig <- rna_effects_all %>%
  dplyr::group_by(Variant) %>%
  dplyr::summarise(
    Total       = dplyr::n(),
    Significant = sum(p.value < pval_cutoff, na.rm = TRUE),
    Pct_Sig     = round(100 * Significant / Total, 1),
    Mean_Abs_Effect = round(mean(abs(value), na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  dplyr::filter(!grepl("Intercept", Variant)) %>%
  dplyr::arrange(dplyr::desc(Significant))

lm_sig %>%
  gt::gt() %>%
  gt::tab_header(
    title    = gt::md("**Linear Model Summary**"),
    subtitle = gt::md(paste0("Significance: p < ", pval_cutoff))
  ) %>%
  gt::data_color(columns = Pct_Sig, method = "numeric",
                 palette = c("white", "#164B6E"))
```

---

## 4. AMP-AD Module Correlation — LM Estimates {.tabset .tabset-pills}

Pearson correlation between per-gene LM variant estimates and human AMP-AD
module fold-changes.

```{r module_lm_check}
if (!run_module_corr_lm || !run_lm_analysis) {
  cat("LM module correlation skipped.\n",
      "Set run_module_corr_lm <- TRUE and run_lm_analysis <- TRUE.\n")
}
```

### 4.1 All Samples

```{r lm_corr_all, fig.width=20, fig.height=7,
    eval=run_module_corr_lm && run_lm_analysis}

lm_cor_all <- corr_function_lm(
  data        = rna_effects_all %>% dplyr::select(Gene, Variant, value),
  human_data  = ampad_modules_fc,
  module_meta = module_clusters,
  mod_order   = mod,
  var_order   = order_variant_lm
)

# Add group column for faceting
lm_cor_all <- lm_cor_all %>%
  dplyr::mutate(
    Background = dplyr::recode(as.character(Variant),
                               !!!lm_variant_groups, .default = "Other"),
    Background = factor(Background, levels = lm_group_order)
  )

if (nrow(lm_cor_all) > 0) {
  ran <- ceiling(max(abs(range(lm_cor_all$correlation))) * 100) / 100
  variant_corrplot_ggh4x(lm_cor_all, ran)
} else {
  message("No LM-module correlations found.")
}
```

### 4.2 By Age Group

```{r lm_corr_age, fig.width=20, fig.height=6,
    eval=run_module_corr_lm && run_lm_analysis}
for (age_label in names(rna_effects_by_age)) {
  lm_cor_age <- corr_function_lm(
    data        = rna_effects_by_age[[age_label]] %>%
                  dplyr::select(Gene, Variant, value),
    human_data  = ampad_modules_fc,
    module_meta = module_clusters,
    mod_order   = mod,
    var_order   = order_variant_lm
  ) %>%
    dplyr::mutate(
      Background = dplyr::recode(as.character(Variant),
                                 !!!lm_variant_groups, .default = "Other"),
      Background = factor(Background, levels = lm_group_order)
    )

  if (nrow(lm_cor_age) > 0) {
    ran <- ceiling(max(abs(range(lm_cor_age$correlation))) * 100) / 100
    print(variant_corrplot_ggh4x(lm_cor_age, ran) +
            ggplot2::ggtitle(paste("LM Module Correlation —", age_label)))
  }
}
```

---

## 5. AD Subtype Correlation {.tabset .tabset-pills}

Correlation between mouse DESeq2 logFC values and human AD subtype signatures
from Nikhil et al. (ROSMAP, Mayo, MSBB) and Neff et al. (MSBB).

```{r subtype_check}
if (!run_subtype_corr) {
  cat("AD subtype correlation skipped.\n",
      "Set run_subtype_corr <- TRUE and provide subtype data files.\n")
}
```

### 5.1 Nikhil Subtypes — logFC

```{r subtype_nikhil, fig.width=9, fig.height=8, eval=run_subtype_corr}
order_all_models <- c(order_model_female, order_model_male)

res_nikhil <- run_subtype_correlation(
  load2_fc        = res_deseq$df,
  target_data     = ampad_subtype_fc,
  cohort_data     = module_cohort,
  subtypes_levels = subtypes,
  sex_levels      = order_all_models
)

if (!is.null(res_nikhil) && nrow(res_nikhil) > 0) {
  ran <- ceiling(max(abs(range(res_nikhil$correlation))) * 100) / 100
  subtype_FC_corrplot_colored(res_nikhil, ran)
} else {
  message("No Nikhil subtype correlations computed.")
}
```

### 5.2 Neff Subtypes — logFC

```{r subtype_neff, fig.width=8, fig.height=8, eval=run_subtype_corr}
res_neff <- run_subtype_correlation(
  load2_fc     = res_deseq$df,
  target_data  = neff_subtype_fc,
  sex_levels   = order_all_models,
  cohort_data  = NULL,
  custom_label = "MSBB_PHG_Neff"
)

if (!is.null(res_neff) && nrow(res_neff) > 0) {
  ran <- ceiling(max(abs(range(res_neff$correlation))) * 100) / 100
  subtype_FC_corrplot_colored(res_neff, ran)
} else {
  message("No Neff subtype correlations computed.")
}
```

---

## 6. BioDomain Correlation — LM Estimates {.tabset .tabset-pills}

Correlation between per-gene LM variant estimates and AMP-AD biological domain
reference fold-changes (biodom.data.FC and subbiodom.data.FC).

```{r biodomain_lm_check}
if (!run_lm_analysis || !run_biodomain_ora) {
  cat("BioDomain LM correlation skipped.\n",
      "Set run_lm_analysis <- TRUE and run_biodomain_ora <- TRUE.\n")
}
```

### 6.1 BioDomain — All Samples

```{r biodomain_all, fig.width=16, fig.height=7,
    eval=run_lm_analysis && run_biodomain_ora}

domain_order <- c(
  "Apoptosis", "APP Metabolism", "Autophagy", "Cell Cycle", "DNA Repair",
  "Endolysosome", "Epigenetic", "Immune Response", "Lipid Metabolism",
  "Metal Binding and Homeostasis", "Mitochondrial Metabolism",
  "Myelination", "Oxidative Stress", "Proteostasis", "RNA Spliceosome",
  "Structural Stabilization", "Synapse", "Tau Homeostasis", "Vasculature"
)

df_lm <- rna_effects_all %>% dplyr::select(Gene, Variant, value)

bd_all <- corr_function_lm_biodomain(
  data           = df_lm,
  ref            = biodom.data.FC,
  sig            = pval_cutoff,
  age_group_val  = "All Samples",
  variant_levels = order_variant_lm,
  domain_levels  = domain_order
) %>%
  dplyr::mutate(
    group = dplyr::recode(as.character(Variant),
                          !!!lm_variant_groups, .default = "Other"),
    group = factor(group, levels = lm_group_order),
    Name  = "Biological Domains"
  )

if (nrow(bd_all) > 0) {
  ran <- ceiling(max(abs(range(bd_all$correlation))) * 100) / 100
  variant_corrplot_biodom(bd_all, ran)
} else {
  message("No BioDomain LM correlations found.")
}
```

### 6.2 Sub-BioDomain — All Samples

```{r sub_biodomain_all, fig.width=12, fig.height=40,
    eval=run_lm_analysis && run_biodomain_ora}

sub_bd_all <- corr_function_lm_subbiodomain(
  data           = df_lm,
  ref            = subbiodom.data.FC,
  sig            = pval_cutoff,
  age_group_val  = "All Samples",
  variant_levels = order_variant_lm,
  domain_levels  = domain_order
) %>%
  dplyr::mutate(
    group = dplyr::recode(as.character(Variant),
                          !!!lm_variant_groups, .default = "Other"),
    group = factor(group, levels = lm_group_order),
    brain = "Biodomains_Treat-AD"
  )

if (nrow(sub_bd_all) > 0) {
  variant_corrplot_subbiodom_portrait(sub_bd_all, ran = 1)
} else {
  message("No Sub-BioDomain LM correlations found.")
}
```

### 6.3 Sub-BioDomain — By Age Group

```{r sub_biodomain_age, fig.width=16, fig.height=7,
    eval=run_lm_analysis && run_biodomain_ora}
for (age_label in names(rna_effects_by_age)) {
  df_age <- rna_effects_by_age[[age_label]] %>% dplyr::select(Gene, Variant, value)

  sub_bd_age <- corr_function_lm_subbiodomain(
    data           = df_age,
    ref            = subbiodom.data.FC,
    sig            = pval_cutoff,
    age_group_val  = age_label,
    variant_levels = order_variant_lm,
    domain_levels  = domain_order
  ) %>%
    dplyr::mutate(
      group = dplyr::recode(as.character(Variant),
                            !!!lm_variant_groups, .default = "Other"),
      group = factor(group, levels = lm_group_order)
    )

  if (nrow(sub_bd_age) > 0) {
    print(variant_corrplot_subbiodom_portrait(sub_bd_age, ran = 1) +
            ggplot2::ggtitle(paste("Sub-BioDomain —", age_label)))
  }
}
```

---

## 7. GO — BioDomain ORA {.tabset .tabset-pills}

GO over-representation analysis for differentially expressed genes, with
results annotated to AMP-AD biological domains and sub-domains.

```{r ora_check}
if (!run_biodomain_ora) {
  cat("BioDomain ORA skipped.\n",
      "Set run_biodomain_ora <- TRUE and provide biodom_path.\n")
}
```

```{r ora_run, eval=FALSE}
# Run ORA — this is computationally intensive, run once and save
univ_ids <- unique(as.data.frame(org.Mm.egGO)$gene_id)
univ_mapped <- clusterProfiler::bitr(
  univ_ids, fromType = "ENTREZID", toType = "SYMBOL",
  OrgDb = org.Mm.eg.db, drop = TRUE
)
univ <- unique(univ_mapped$SYMBOL)

enr_ora <- run_enrichment_pipeline(res_deseq$df, univ)
save(enr_ora, file = file.path(result_dir, "ORA_BioDomain_RNAseq.RData"))
cat("ORA results saved.\n")
```

```{r ora_load, eval=run_biodomain_ora}
load(file.path(result_dir, "ORA_BioDomain_RNAseq.RData"))
```

### 7.1 BioDomain ORA — All Models

```{r ora_plot, fig.width=12, fig.height=10, eval=run_biodomain_ora}
enr_ora_plot <- enr_ora %>%
  dplyr::filter(!(Biodomain %in% "none")) %>%
  dplyr::mutate(model = factor(model, levels = rev(unique(model))))

if (nrow(enr_ora_plot) > 0) {
  ORA_plot_fn(enr_ora_plot)
} else {
  message("No BioDomain ORA results to plot.")
}
```

### 7.2 Sub-BioDomain ORA

```{r sub_ora_plot, fig.width=20, fig.height=22, eval=run_biodomain_ora}
top_bd <- enr_ora_plot %>%
  dplyr::filter(Biodomain != "none") %>%
  dplyr::count(Biodomain, sort = TRUE) %>%
  head(top_biodomains) %>%
  dplyr::pull(Biodomain)

sub_enr <- enr_ora_plot %>%
  dplyr::filter(Biodomain %in% top_bd)

if (nrow(sub_enr) > 0) {
  plot_biodomain_distribution(sub_enr,
    paste("Sub-BioDomain ORA — Top", top_biodomains, "Domains"))
} else {
  message("No sub-domain ORA results to plot.")
}
```

---

## 8. GO — BioDomain GSEA {.tabset .tabset-pills}

GSEA using all genes ranked by log2 fold-change, annotated to AMP-AD
biological domains. Unlike ORA, does not require a significance cutoff.

```{r gsea_check}
if (!run_biodomain_gsea) {
  cat("BioDomain GSEA skipped.\n",
      "Set run_biodomain_gsea <- TRUE and provide biodom_path.\n")
}
```

```{r gsea_run, eval=FALSE}
enr_gsea <- enrichGSEA_BD(
  dat  = res_deseq$df,
  set1 = unique(res_deseq$df$group)
)
save(enr_gsea, file = file.path(result_dir, "GSEA_BioDomain_RNAseq.RData"))
cat("GSEA results saved.\n")
```

```{r gsea_load, eval=run_biodomain_gsea}
load(file.path(result_dir, "GSEA_BioDomain_RNAseq.RData"))
```

### 8.1 BioDomain GSEA — Significant Terms

```{r gsea_plot, fig.width=12, fig.height=9, eval=run_biodomain_gsea}
gsea_sig <- enr_gsea %>%
  dplyr::filter(p.adjust < 0.1, !(Biodomain %in% "none"))

if (nrow(gsea_sig) > 0) {
  plot_gsea_biodomain_distribution(gsea_sig, ran = nes_axis_limit)
} else {
  message("No significant GSEA terms (p.adjust < 0.1).")
}
```

### 8.2 Sub-BioDomain GSEA — Females

```{r gsea_sub_female, fig.width=16, fig.height=7, eval=run_biodomain_gsea}
gsea_f <- enr_gsea %>%
  dplyr::filter(grepl("Female", model), p.adjust < 0.4,
                !(Biodomain %in% "none"))

if (nrow(gsea_f) > 0) {
  ggplot2::ggplot(gsea_f, ggplot2::aes(NES, Subdomain)) +
    ggplot2::geom_violin(data = subset(gsea_f, NES > 0),
                         ggplot2::aes(color = color), scale = "width") +
    ggplot2::geom_violin(data = subset(gsea_f, NES < 0),
                         ggplot2::aes(color = color), scale = "width") +
    ggplot2::geom_jitter(ggplot2::aes(size = -log10(p.adjust), fill = color),
                         color = "grey20", shape = 21, alpha = 0.3) +
    ggplot2::geom_vline(xintercept = 0, lwd = 0.5, lty = 2) +
    ggplot2::scale_y_discrete(drop = FALSE) +
    ggplot2::scale_fill_identity() + ggplot2::scale_color_identity() +
    ggplot2::ggtitle("Sub-BioDomain GSEA — Females") +
    ggplot2::facet_grid(rows = dplyr::vars(abbr), cols = dplyr::vars(model),
                        scales = "free", space = "free", switch = "y") +
    ggplot2::theme_bw() +
    ggplot2::theme(
      strip.text.y.left = ggplot2::element_text(angle = 0, size = 12),
      axis.text.x       = ggplot2::element_text(angle = 0, hjust = 0, size = 11),
      axis.text.y       = ggplot2::element_text(size = 11)
    )
} else {
  message("No female GSEA sub-domain results at p.adjust < 0.4.")
}
```

### 8.3 Sub-BioDomain GSEA — Males

```{r gsea_sub_male, fig.width=16, fig.height=10, eval=run_biodomain_gsea}
gsea_m <- enr_gsea %>%
  dplyr::filter(grepl("Male", model), p.adjust < 0.1,
                !(Biodomain %in% "none"))

if (nrow(gsea_m) > 0) {
  ggplot2::ggplot(gsea_m, ggplot2::aes(NES, Subdomain)) +
    ggplot2::geom_violin(data = subset(gsea_m, NES > 0),
                         ggplot2::aes(color = color), scale = "width") +
    ggplot2::geom_violin(data = subset(gsea_m, NES < 0),
                         ggplot2::aes(color = color), scale = "width") +
    ggplot2::geom_jitter(ggplot2::aes(size = -log10(p.adjust), fill = color),
                         color = "grey20", shape = 21, alpha = 0.3) +
    ggplot2::geom_vline(xintercept = 0, lwd = 0.5, lty = 2) +
    ggplot2::scale_y_discrete(drop = FALSE) +
    ggplot2::scale_fill_identity() + ggplot2::scale_color_identity() +
    ggplot2::ggtitle("Sub-BioDomain GSEA — Males") +
    ggplot2::facet_grid(rows = dplyr::vars(abbr), cols = dplyr::vars(model),
                        scales = "free", space = "free", switch = "y") +
    ggplot2::theme_bw() +
    ggplot2::theme(
      strip.text.y.left = ggplot2::element_text(angle = 0, size = 12),
      axis.text.x       = ggplot2::element_text(angle = 0, hjust = 0, size = 11),
      axis.text.y       = ggplot2::element_text(size = 11)
    )
} else {
  message("No male GSEA sub-domain results at p.adjust < 0.1.")
}
```

---

## 9. Save Results

```{r save_all}
results_to_save <- list()

if (run_module_corr_fc) {
  results_to_save$module_corr_female_fc <- corr_res_female
  results_to_save$module_corr_male_fc   <- corr_res_male
}
if (run_module_corr_lm && run_lm_analysis) {
  results_to_save$lm_module_corr_all    <- lm_cor_all
}
if (run_subtype_corr) {
  results_to_save$subtype_nikhil        <- res_nikhil
  results_to_save$subtype_neff          <- res_neff
}
if (run_lm_analysis) {
  results_to_save$lm_effects_all        <- rna_effects_all
  results_to_save$lm_effects_by_age     <- rna_effects_by_age
}
if (run_biodomain_ora) {
  results_to_save$ora_biodomain         <- enr_ora
  if (exists("bd_all"))     results_to_save$lm_biodomain_all     <- bd_all
  if (exists("sub_bd_all")) results_to_save$lm_subbiodomain_all  <- sub_bd_all
}
if (run_biodomain_gsea) {
  results_to_save$gsea_biodomain        <- enr_gsea
}

if (length(results_to_save) > 0) {
  save(results_to_save,
       file = file.path(result_dir, "Human_Translation_RNAseq.RData"))
  cat("Results saved to:",
      file.path(result_dir, "Human_Translation_RNAseq.RData"), "\n")
  cat("Objects saved:\n")
  cat(paste0("  - ", names(results_to_save), collapse = "\n"), "\n")
} else {
  message("No results to save — all switches are FALSE.")
}
```

---

## Session Info

```{r session_info, echo=TRUE}
sessionInfo()
```
